---
title: "Practical Machine Learning Final Project-Report"
author: "John Kennedy Mwangi"
date: "February 2, 2019"
output: html_document
---
#Loading the data from csv's and load required libraries

```{r}
traindat <- read.csv('./pml-training.csv', header=T)
testdat <- read.csv('./pml-testing.csv', header=T)

library(caret)
library(rattle)

```
#Check the dimensions and the structure of the loaded data

```{r}
dim(traindat)
str(traindat)
dim(testdat)
str(traindat)
```
# Remove the first five variables that have no apparent weight on the varibale "classe"
```{r}
traindat <- traindat[,-c(1,2,3,4,5)]
testdat <- testdat[,-c(1,2,3,4,5)]
```

#Get rid of all columns with non zero variance based on the training data
```{r}
nonzerovariance <- nearZeroVar(traindat)
traindat <- traindat[,-nonzerovariance]
testdat <- testdat[,-nonzerovariance]
```

# Get rid of varibales and columns with NAs
```{r}
removeNA <- apply(traindat, 2, function(x) { sum(is.na(x)) } )
traindat <- traindat[,which(removeNA ==0)]
testdat <- testdat[,which(removeNA ==0)]
```
# Get rid of highly correlated variables
```{r}
HighCor <- findCorrelation(cor(traindat[, 1:dim(traindat)[2]-1]), cutoff=0.8)
traindat <- traindat[,-HighCor]
testdat <- testdat[,-HighCor]
```
# Use complete.case function to keep the data that has all field values
```{r}
traindat <- traindat[complete.cases(traindat),]
testdat <- testdat[complete.cases(testdat),]
dim(traindat)
str(traindat)
dim(testdat)
str(testdat)

```
# Partition the training data into two datsets for training and testing the models
```{r}
set.seed(12345)
trainPartition <- createDataPartition(traindat$classe, p = 0.7, list = FALSE)
Traindata1  <- traindat[trainPartition, ]
Traindata2 <- traindat[-trainPartition, ]

```
### Use Traindata1 to create models with Random Forest and Classification Trees
```{r}
library(randomForest)
controlRandom <- trainControl(method="none", number=3, verboseIter=FALSE)

RandomModel <- train(classe~., data=Traindata1,  method="rf", trControl=controlRandom, verbose=FALSE)

```
```{r}
library(rpart)
Classificationmodel <- train(classe~., data=Traindata1, method="rpart", trControl=controlRandom )
```

### Test the models on Traindata2
```{r}
library(randomForest)
predictTrainRand <- predict(RandomModel, newdata=Traindata2)



```
```{r}
library(rpart)

predictTrainClass <- predict(Classificationmodel, newdata=Traindata2)
```

### Obtain confusionmatrices for the two models and compare
```{r}
library(randomForest)
confMatRandom <- confusionMatrix(predictTrainRand, Traindata2$classe)
confMatRandom




```
```{r}
library(rpart)

confMatClass <- confusionMatrix(predictTrainClass, Traindata2$classe)
confMatClass 
```

#Test the random forest model on the initial test data 
```{r}
library(randomForest)
predictTestData <- predict(RandomModel, newdata=testdat)
predictTestData
```


